
% Carl-Johan Hoel
% Ivo Batkovic
% Arian Ranjbar
% Maxime Bouton 
% Anton och robin, Started as a master thesis student ended up as friends

% WASP 
% friends and colleagues Krook, John, Daniel, Dapeng, Erik, Emil, Anton, Rui, Lars, Linnea, Oskar, Christian, Viktor, and the rest of my colleagues from Chalmers, Zenseact, and WASP.

% SISL 
% Mykel Kochenderfer
% Mattias, shushman, Jayesh, Jeremy, Ayan 
% Wallenberg stanford 

This thesis is the result of many years of hard work, lots of travel and a great learning experience. Even in times when things did not work as expected or unexpected problems emerged, I never felt that I was alone. That is why I would like to acknowledge the people who have made this work possible and memorable. 

First, I want to thank my supervisor and examiner Professor Jonas Sj\"oberg. 
% We have had our ups and downs but in the end I respect you and the achievements you have accomplished. 
Thank you for believing in me and the oppertunity of pursuing a doctoral degree as your student. It has been an honor and a pleasure.

A special thanks to my industrial supervisor Dr.~Mohammad Ali, who chose and encouraged me to pursue a PhD in this field. I still remember the pitch that convinced me to do it. You said: "Pilot assist is mostly done now, we need someone to look towards the future of decision making and control". We had a lot of fun in the beginning of this journey and I will always be grateful for the guidance and support you provided, and most of all for believing in me. 

The work presented in this thesis would not have been possible without the financial support from VCC, Zenuity, Zenseact, and the Wallenberg AI, Autonomous Systems and Software Program~(WASP). I am also grateful for the community and environment created by both Zenseact and WASP. The Advanced Graduate Program led by Dr. Mats Nordlund and Dr. Carl Lindberg created an environment within the company created a safe space to share ideas, talk amongst peers and fun to go to work in the morning. 

I would especially like to thank Dr. Carl Lindberg for the coaching during these final years of my PhD. At the start of the pandemic, you paid attention to our casual conversations, identified problems and proposed solutions I did not think was possible, but more importantly you cared. Thanks to you, I was able to finish my final year of my PhD together with my wife on the other side of the world and as a direct result of that, I now have a beautiful son. For this, I am eternally grateful. 
 
I especially want to thank my great friends and co-authors with whom I embarked on this PhD journey with. Carl-Johan Hoel, my research twin, thank you for all the support over the years, the enlightening conversations and most of all for being a friend. %I can confidently say that 
Ivo Batkovic, thank you for all the good times we shared. I have always admired your determination and focus working with you has been a pleasure and a privilege. I am happy to call you my friend. 
I would also like to thank Anton Jansson and Robin Gr\"onberg for their well executed master thesis from where I found the direction of this thesis. We started out as supervisor and master thesis students but ended up as friends. 
\todo{Hannes and debba}
Special thanks go to Christian Rodriguez for being a great friend and supporting me whenever times were tough.

I would like to thank WASP for all the interesting study trips, courses, and events that not only widened my skills, but also introduced me to a vast network of colleagues and friends across the world. In addition, the WASP exchange program gave me the opportunity to spend six months as a visiting research student at the Stanford Intelligent Systems Laboratory. To that end, I am deeply grateful that Professor Mykel Kochenderfer that thought me that life is a POMDP and gave me the opportunity to spend time and work with his research group. 
Additionally, special thanks to Maxime Bouton 
\todo{continue}

% I especially want to thank my great friends with whom I started the PhD journey with. Arian Ranjbar, you have been a great office-mate, and you always ensured that it was fun to work from ``skrubben'' (our Chalmers ``office''). Tommy Tram, you have been a great colleague and a great friend throughout this journey. I will always be grateful for your endless support and for all the good times we have shared. I am sure there are more to come. Ankit Gupta, it was impossible to walk past your office without starting some sort of conversation with you. I'm not sure how many hours we spent talking, instead of doing actual research. Thank you for all the good times. Special thanks go to Amrit Krishnan and Fredrik Hoxell for being great friends, and supporting me whenever times were tough. I am also deeply grateful to my friends and colleagues Giuseppe, Robert, Stefan, Krook, John, Daniel, Dapeng, Erik, Emil, Anton, Rui, Lars, Ramin, Simon, Fredrik, Elena, Masoud, Angelos, Angel, Remi, Albert, Carl-Johan, Muhammad, Sten, Maximilian, and the rest of my colleagues from Chalmers, Zenseact, and WASP.


% I would like to thank WASP for all the interesting study trips, courses, and events that not only widened my skills, but also introduced me to a vast network of colleagues and friends across the world. In addition, the WASP exchange program gave me the opportunity to spend six months as a visiting PhD student at the Model Predictive Control Lab at U.C. Berkeley. To that end, I am deeply grateful that Professor Francesco Borrelli gave me the opportunity to spend time and work with his research group. Additionally, special thanks go to the MPC lab, especially to Vijay, Xu, Tommaso, Ugo, and Monimoy, for making my exchange a pleasant and memorable experience.

% Na kraju bih želio posebno zahvaliti svojoj obitelji, ocu Zvonki, majci Marici, sestri Antoniji i bratu Josipu. Osobitu zahvalnost dugujem zaručnici Ivani. Hvala vam svima na beskrajnoj ljubavi i podršci tijekom ovog dijela mog života. Bez vas ovo ne bi bilo moguće.


\hspace*{0pt}\hfill\emph{Tommy Tram,\\\hspace*{0pt}\hfill Göteborg, \TheMonth, \TheYear}.



